\section{Details of the Algorithm}
\label{sec:description}


\input{localRecovery/pseudocode-localrecovery}

A pseudo-code of the local recovery algorithm is given in \pseudosref{local-data|structures}{local-seek:successor}.
The pseudo-code only shows the seek phase of an algorithm and not its \action{} phase since the \action{} phase is algorithm-specific. 

The local recovery algorithm assumes that the original algorithm supports the following functions:
\begin{enumerate*}[label=(\alph*)]
\item \GetKey(~), \IsMarked(~) and \GetChild(~) return the various attributes of a tree node,
\item \IsNull(~) returns true if a reference is null and false otherwise,
\item \GetAddress(~) returns the node address stored in a reference, if non-null,
\item \Move(~) enables the original algorithm to move along an edge, which may invoke helping and cause traversal to restart as in~\cite{HowJon:2012:SPAA},
\item \NeedCleanParentNode(~) returns true if the operation needs the parent node to be clean and have no operation in progress (needed for a delete operation since it needs to modify a child pointer at the parent node), and
\item \PopulateSeekRecord(~) copies the relevant information from the traversal state required by the algorithm into a seek record (\emph{e.g.}, addresses of the terminal node and its parent).
\end{enumerate*}

\Pseudoref{local-data|structures} shows the data structures used by the local recovery algorithm.
Note that all the data structures shown in \pseudoref{local-data|structures} are \emph{local} to a process not shared among processes.
A process uses three main data structures, namely \TraversalRecord{}, \OpRecord{} and \SeekRecord{}. 
A \TraversalRecord{} (\linesref{local-traversal|record:begin}{local-traversal|record:end}) is essentially a stack used to store the nodes visited during tree traversal when looking for a key.
Each entry in a traversal stack (\linesref{local-stack|entry:begin}{local-stack|entry:end}) stores the address of the node, the location of its closest \myanchor{} node (within the stack) and whether the node is a left or right child of its parent.
An \OpRecord{} (\linesref{local-op|record:begin}{local-op|record:end}) stores information about the operation such as type and key as well two stacks: one used when looking for the target key (all operations) and one used when looking for the successor key (only complex delete operations).
Finally, a \SeekRecord{} (\linesref{local-seek|record:begin}{local-seek|record:end}) is used to return the outcome of a tree traversal to the original algorithm.
Its fields are algorithm-specific. 
%%
For example, for \CASTLE{}, \SeekRecord{} contains three fields: 
\begin{enumerate*}[label=(\alph*)]
\item two addresses, namely those of the target node and its parent, and
\item the contents of the injection point where an insert operation needs to attach the new node. 
\end{enumerate*}

\Pseudoref{local-stack|functions} shows the functions used to manipulate a traversal stack. 
The function \Size{} (\linesref{local-size:begin}{local-size:end}) returns the number of entries in the stack. 
The functions \GetTop{} (\linesref{local-get|top:begin}{local-get|top:end}) and \GetSecondToTop{} (\linesref{local-get|second|to|top:begin}{local-get|second|to|top:end}) return the address of the node stored in the topmost entry and the entry below it, respectively. 
The function \AddToTop{} (\linesref{local-add|to|top:begin}{local-add|to|top:end}) adds an entry to the top of the stack while \RemoveFromTop{} (\linesref{local-remove|from|top:begin}{local-remove|from|top:end}) removes an entry from the top of the stack. 
The function \RemoveUntilCritical{} (\linesref{local-remove|until|critical:begin}{local-remove|until|critical:end}) removes the entries from the top of the stack until a given point.
The function \RememberCritical{} (\linesref{local-remember|critical:begin}{local-remember|critical:end}) updates the \myanchor{} field of the \myanchor{} node of the topmost entry in the stack.
The function \GetFullEntry{} (\linesref{local-get|full|entry:begin}{local-get|full|entry:end} returns all the three fields of a given entry in the stack (may not be the topmost entry). 
The function \InitializeTraversalRecord{} (\linesref{local-initialize|traversal|record:begin}{local-initialize|traversal|record:end}) initializes a traversal stack using sentinel nodes.

\Pseudosref[ \& ]{local-local:recovery}{local-seek} show the functions used to look for a key in the tree.
The function \SeekForTarget{} (\linesref{local-seek:begin}{local-seek:end}) first backtracks to a \mysafe{} node in the stack (\lineref{local-seek:while:find|start|point}).
Initially, the start point of the traversal will be a sentinel node, which is always a \mysafe{} node.
The function then traverses the tree from top to down by following either the left or the right child pointer (\lineref{local-seek:while:traversal:select}) until it either finds the key or encounters a null pointer (\linesref{local-seek:while:traversal:stop:begin}{local-seek:while:traversal:stop:end}).
In case the terminal node of the traversal does not contain the matching key, the function tests whether or not the terminal node is \mysafe{} by invoking  \FindAdmissible{} function (\lineref{local-seek:while:traversal:test|safety}).
If the node is found to be \mysafe{}, then the traversal is ended (\lineref{local-seek:while:traversal:safe}.
Otherwise, the traversal path would have been truncated in which case the function determines if the traversal can still stop by invoking \ExamineTop{} function that looks at the last node in the path (\lineref{local-seek:while:traversal:not|safe}). 
As the traversal moves down the tree, the function also populates the traversal stack (\lineref{local-seek:while:traversal:move}).
The function \FindAdmissible{} (\linesref{local-test|safety:begin}{local-test|safety:end}) implements the method described in \secref{test|safety}.
The function \FindStartPoint{} (\linesref{local-find|start|point:begin}{local-find|start|point:end}) implements the method described in \secref{find|restart|point}.
The function \ExamineTop{} (\linesref{local-examine|top:begin}{local-examine|top:end}) examines the topmost node in the stack. If the node is found to be a \myinconsistent{} node, then the traversal can stop for a search and delete operation (\lineref{local-examine|top:inconsistent}).
If the node's key matches the operation's key, then also the traversal can stop (\lineref{local-examine|top:matching}).
Otherwise, the traversal needs to be restarted (\linesref[ \& ]{local-examine|top:continue|1}{local-examine|top:continue|2}).

\Pseudoref{local-seek:successor} shows the function \SeekForSuccessor{} used to locate the successor key by a complex delete operation (\linesref{local-seek|successor:begin}{local-seek|successor:end}). The function first backtracks to an unmarked node with a clean parent if required (\linesref{local-seek|successor:while:backtrack:begin}{local-seek|successor:while:clean:end}). It then checks whether or not the successor key is still needed by invoking \NeedSuccessorKey{} function (\lineref{local-seek|successor:while:need|successor}). The function \NeedSuccessorKey{} returns a reference, which is null if the successor key is no longer needed and contains the address of the target node's right child otherwise. This address is used as a traversal point if the stack only contains a single entry (the node whose key needs to be replaced). If the successor key is still needed, then the function repeatedly follows the left child pointer until it encounters a null pointer (\linesref{local-seek|successor:while:traversal:begin}{local-seek|successor:while:traversal:end}). While moving down the tree, the function also populates the traversal stack (\lineref{local-seek|successor:while:traversal:stack}).
\input{localRecovery/pseudocode-seekForSuccessor}