\newenvironment{limitscope}{}{}
\begin{limitscope}
%%%%% castle macros - begin
\newcommand{\accesspath}{access-path}
\newcommand{\terminalnode}{terminal node}

\newcommand{\true}{\textsf{true}}
\newcommand{\false}{\textsf{false}}

\newcommand{\CAS}{\textsf{CAS}}

\newcommand{\sNodeOne}{\mathbb{R}}
\newcommand{\sNodeTwo}{\mathbb{S}}
\newcommand{\sKeyOne}{\infty_1}
\newcommand{\sKeyTwo}{\infty_2}

\newcommand{\targetnode}{target node}
\newcommand{\anchornode}{anchor node}

\newcommand{\myparent}{parent}
\newcommand{\myleft}{le\!f\!t}
\newcommand{\myright}{right}

\newcommand{\CASTLE}{\textsc{CASTLE}}
\newcommand{\CITRUS}{\textsc{CITRUS}}
\newcommand{\HJBST}{\textsc{LF-IBST}}
\newcommand{\NMBST}{\textsc{LF-EBST}}

\newcommand{\RemoveChild}{\textsc{RemoveChild}}
\newcommand{\LockAll}{\textsc{LockAll}}
\newcommand{\UnlockAll}{\textsc{UnlockAll}}
\newcommand{\ClearFlags}{\textsc{ClearFlags}}
\newcommand{\FindSmallest}{\textsc{FindSmallest}}

\newcommand{\lFlag}{lFlag}
\newcommand{\mFlag}{mFlag}
\newcommand{\nFlag}{nFlag}

%%%%% castle macros - end

\section{The Lock-Based Algorithm}
\label{sec:algorithm}
We first provide an overview of our algorithm. We then describe the algorithm in more detail and also give its pseudo-code. For ease of exposition, we describe our algorithm assuming no memory reclamation, which can be performed using the well-known technique of hazard pointers~\cite{Mic:2004:TPDS}.

\input{c03/overview}

\input{c03/description}

\subsection{Formal Description}

We refer to our algorithm as \CASTLE{} (\underline{C}oncurrent \underline{A}lgorithm for Binary \underline{S}earch \underline{T}ree by \underline{L}ocking \underline{E}dges). 

\input{c03/pseudocode}

A pseudo-code of our algorithm is given in \algosref{data|structures}{helper}.
Different data structures used in our algorithm are shown in \algoref{data|structures}. Besides tree node, we use three additional records:
\begin{enumerate*}[label=(\alph*)]
\item  \emph{seek record:} to store the outcome of a tree traversal both when looking for the target key and the successor key, 
\item \emph{anchor record:} to store information about the \anchornode{} during the seek phase, and
\item  \emph{lock record:} to store information about a tree edge that needs to be locked. 
\end{enumerate*}

The pseudo-code for the seek function is shown in \algoref{seek}. The 
pseudo-codes for search, insert and delete operations are shown  in 
\algoref{search}, \algoref{insert} and \algoref{delete}, respectively. 
\Algoref{lock:unlock} contains the pseudo-code for locking and unlocking a 
set of tree edges, as specified in an array. Finally, \algoref{helper} contains the pseudo-codes for 
three helper functions used by a delete operation, namely:
\begin{enumerate*}[label=(\alph*)]
\item \ClearFlags{}: to clear lock and mark flags from a child field, 
\item \FindSmallest{}: to locate the smallest key in a subtree, and
\item \RemoveChild{}: to remove a given child of a node.
\end{enumerate*}

In the pseudo-code, to improve clarity, we sometimes use subscripts $l$, $m$ and $n$ to denote lock, mark and null flags, respectively.  
\input{c03/proof}

\end{limitscope}